---
title: "Capstone - Cleaning & Random Forest"
output: pdf_document
date: "2023-03-26"
---

```{r, message=FALSE, warning=FALSE}
# setup
set.seed(478)
library(dplyr)
library(ggplot2)
library(tidymodels)
library(tidyverse)
library(randomForest)
library(rpart)
library(vip)
library(Boruta)
```


```{r}
# read in data 
data_nfl <- read.csv('https://raw.githubusercontent.com/kristinewiggins/NFL-Pressure-Kicks/c01bd037915638fc425517a2a362aa9b745c903b/NFLkicks.csv')
```

```{r}
glimpse(data_nfl)
```

```{r}
# filter out observations without timeout called
timeout_occures <- data_nfl[data_nfl$timeout > 0,]

# clean up NAs in temp, wind, play_clock,  of default appropriate values
timeout_occures <- timeout_occures %>% mutate(temp = replace_na(temp, 65))
timeout_occures <- timeout_occures %>% mutate(wind = replace_na(wind, 0))
timeout_occures <- timeout_occures %>% mutate(play_clock = replace_na(play_clock, 0))
timeout_occures <- timeout_occures %>% mutate(drive = replace_na(drive, 0))

# Fixing NAs in roof variable by looking up each team's stadium
which(is.na(timeout_occures), arr.ind=TRUE)
timeout_occures[2488, 44] <- "open" #add retractable? 
timeout_occures[2492, 44] <- "dome"
timeout_occures[2498, 44] <- "open" #add retractable? 
timeout_occures[2508, 44] <- "open" #add retractable? 
timeout_occures[2533, 44] <- "open"
timeout_occures[2534, 44] <- "open" 
timeout_occures[2535, 44] <- "open" 
timeout_occures[2549, 44] <- "open" #add retractable? 
timeout_occures[2554, 44] <- "open"
timeout_occures[2555, 44] <- "open"
timeout_occures[2564, 44] <- "open"
timeout_occures[2575, 44] <- "dome" #add retractable? 
timeout_occures[2583, 44] <- "dome" #add retractable? 
timeout_occures[2595, 44] <- "dome" #add retractable? 
timeout_occures[2596, 44] <- "dome" #add retractable? 
timeout_occures[2619, 44] <- "dome" #add retractable?
timeout_occures[2620, 44] <- "dome" #add retractable?
timeout_occures[2628, 44] <- "dome" #add retractable?

sum(is.na(timeout_occures))
```

```{r}
# observations occurring under 60 seconds 
presskicks1 <- timeout_occures[timeout_occures$game_seconds_remaining <= 60,]
```

```{r}
# splitting data 
sample_size = floor(0.7*nrow(presskicks1))
picked = sample(seq_len(nrow(presskicks1)), size = sample_size)
presskicks1_train <- presskicks1[picked,]
presskicks1_test <- presskicks1[-picked,]

# convert categorical binary to numerical binary 
presskicks1_train$field_goal_result <- ifelse(presskicks1_train$field_goal_result == "made",1,0)
presskicks1_test$field_goal_result <- ifelse(presskicks1_test$field_goal_result == "made",1,0)
``` 


```{r}
# random forest 
rforest <- randomForest(as.factor(field_goal_result) ~ ., 
                   data = presskicks1_train, 
                   ntree = 200)
rforest
```

TP + TN / TP + TN + FP + FN 
84 + 279 / 84 + 279 + 0 + 3 = 0.991803278689
99% ??!??! 

```{r}
rforest_import <- randomForest(as.factor(field_goal_result) ~ ., 
                   data = presskicks1_train, 
                   ntree = 200, importance = TRUE)
rforest_import
plot(rforest_import) 
vip(rforest_import)
```
Same accuracy and OOB estimate. 

------------------
Variables in order of importance with description. Importance is a measure of 
by how much removing a variable decreases accuracy, and vice versa — by how much 
including a variable increases accuracy.
------------------
sp:	Binary indicator for whether or not a score occurred on the play.

score_differential_post: Score differential between the posteam and defteam at the end of the play.

def_wp:	Estimated win probability for the defteam.

wp:	Estimated win probability for the posteam given the current situation at the start of the given play.

score_differential: Score differential between the posteam and defteam at the start of the play.

no_score_prob: Predicted probability of no score occurring for the rest of the half based on the expected points model.

kick_distance:	Numeric distance in yards for kickoffs, field goals, and punts.

home_wp:	Estimated win probability for the home team.

temp: The temperature at the stadium only for 'roof' = 'outdoors' or 'open'. 

away_wp:	Estimated win probability for the away team.
-----------------

```{r}
# Removing sp variable as it is redundant
presskicks1_train <- presskicks1_train[, -14]

rforest_import <- randomForest(as.factor(field_goal_result) ~ ., 
                   data = presskicks1_train, 
                   ntree = 200, importance = TRUE)
rforest_import
plot(rforest_import) 
vip(rforest_import)
```

TP + TN / TP + TN + FP + FN 
69 + 276 / 69 + 276 + 18 + 3 = 0.94262295082 
94.26 % accuracy for prediction. 

------------------
Variables in order of importance with description. Importance is a measure of 
by how much removing a variable decreases accuracy, and vice versa — by how much 
including a variable increases accuracy.
------------------
score_differential_post: Score differential between the posteam and defteam at the end of the play.

def_wp:	Estimated win probability for the defteam.

wp:	Estimated win probability for the posteam given the current situation at the start of the given play.

score_differential: Score differential between the posteam and defteam at the start of the play.

no_score_prob: Predicted probability of no score occurring for the rest of the half based on the expected points model.

kick_distance:	Numeric distance in yards for kickoffs, field goals, and punts.

result: Estimated win probability for the home team.

quarter_seconds_remaining: The temperature at the stadium only for 'roof' = 'outdoors' or 'open'. 

away_wp: Estimated win probability for the away team.

home_wp: Estimated win probability for the home team.
-----------------

```{r, message=FALSE, warning=FALSE}
boruta <- Boruta(as.factor(field_goal_result) ~ ., 
                   data = presskicks1_train, doTrace = 2, maxRuns = 500)

print(boruta$finalDecision)
par(mar = c(8,5,2,2))
plot(boruta, las = 2, cex.axis = 0.7)
```

Boruta is a feature selection algorithm. Precisely, it works as a wrapper 
algorithm around Random Forest. This package derive its name from a demon in 
Slavic mythology who dwelled in pine forests. 

```{r}
# running RF on test 
rforest_import <- randomForest(as.factor(field_goal_result) ~ ., 
                   data = presskicks1_test, 
                   ntree = 200, importance = TRUE)
rforest_import
plot(rforest_import) 
vip(rforest_import)

# running RF w/o sp 
presskicks1_test <- presskicks1_test[, -14]
rforest_import <- randomForest(as.factor(field_goal_result) ~ ., 
                   data = presskicks1_test, 
                   ntree = 200, importance = TRUE)
rforest_import
plot(rforest_import) 
vip(rforest_import)

```
27 + 130 / 0 + 1 + 27 + 130 = 0.993670886076
99.37% accurcy 
5 + 130 / 5 + 130 + 23 + 0 = 0.854430379747
85.44% accurcy 

```{r}
#presskicks3 <- timeout_occures[timeout_occures$game_seconds_remaining <= 180,]
#presskicks5 <- timeout_occures[timeout_occures$game_seconds_remaining <= 300,]

```
