---
title: 'NFL Pressure Kicks and Their Correlates'
author: 'Marissa Ellingson, Sophie Giacobbe, Jason Nero, and Kristine Wiggins'
output: pdf_document
---

```{r}
# load necessary libraries
library(nflfastR)
library(tidyverse)
library(ggplot2)
library(randomForest)
library(rpart)
library(vip)
library(Boruta)
library(lme4)
library(GGally)
library(knitr)
library(FactoMineR)
library(factoextra)
library(gridExtra)
library(grid)
library(broom)
library(ISLR)
library(yardstick)
library(discrim)
library(kknn)
library(class)
```

```{r}
## code used to retrieve the dataset and create .csv
#
#pbp <- nflfastR::load_pbp(2000:2022)
#pbp <- pbp[!is.na(pbp$play_type),]
#fgto <- pbp[pbp$play_type == 'field_goal' | pbp$play_type == #'no_play',]
## remove NAs and penalties
#fgto <- fgto[!is.na(fgto$timeout),]
#fgto <- fgto[fgto$penalty != 1,]
## add timeouts to field goal rows
#for (i in 1:nrow(fgto)){
#  if (fgto$timeout[i]==1 & fgto$play_type[i] != 'field_goal'){
#    if (fgto$play_type[i+1]=='field_goal' & #fgto$time[i]==fgto$time[i+1]){
#      fgto$timeout[i+1] = 1
#      fgto$timeout_team[i+1] = fgto$timeout_team[i]
#    }
#  }
#}
## keep only the field goal rows
#fg <- fgto[fgto$play_type == 'field_goal',]
#write.csv(fg, 'NFLfieldgoals.csv')
```

```{r}
## select the useful columns to create final dataset .csv
#fg <- read.csv('NFLfieldgoals.csv')
#fg <- fg %>% select(play_id,game_id,home_team,away_team,
#                season_type,week,posteam,posteam_type,
#                quarter_seconds_remaining,
#                half_seconds_remaining,
#                game_seconds_remaining,drive,sp,qtr,
#                time,play_type,field_goal_result,
#                kick_distance,home_timeouts_remaining,
#                away_timeouts_remaining,timeout,
#                timeout_team,posteam_timeouts_remaining,
#                defteam_timeouts_remaining,posteam_score,
#                defteam_score,score_differential,
#                posteam_score_post,defteam_score_post,
#                score_differential_post,no_score_prob,
#                opp_fg_prob,wp,def_wp,home_wp,away_wp,
#                kicker_player_name,kicker_player_id,
#                season,play_clock,location,
#                result,roof,surface,temp,wind)
#write.csv(fg, 'NFLkicks.csv')
```

```{r}
# read and clean data
kicks <- read.csv('https://raw.githubusercontent.com/kristinewiggins/NFL-Pressure-Kicks/c01bd037915638fc425517a2a362aa9b745c903b/NFLkicks.csv')
kicks <- kicks[kicks$field_goal_result != 'blocked',]
kicks <- kicks %>% select(-X,-sp)

kicks$temp[is.na(kicks$temp)] <- 65
kicks$wind[is.na(kicks$wind)] <- 0
kicks$drive[is.na(kicks$drive)] <- 1
kicks$play_clock[is.na(kicks$play_clock)] <- 0
kicks$roof[is.na(kicks$roof)] <- 'retractable'

presskicks1 <- kicks[kicks$game_seconds_remaining <= 60,]
presskicks3 <- kicks[kicks$game_seconds_remaining <= 180,]
presskicks5 <- kicks[kicks$game_seconds_remaining <= 300,]
```

## FAMD Exploration

```{r}
df <- subset(kicks, select = -c(play_id, game_id, 
                                  quarter_seconds_remaining, 
                                  half_seconds_remaining, time, 
                                  kicker_player_name, home_team, away_team, 
                                  posteam, timeout_team, play_type))
```

```{r}
res.famd <- FAMD(df, graph = FALSE)
print(res.famd)
```

```{r}
#proportion of variances retained by the different dimensions (axes)
eig.val <- get_eigenvalue(res.famd)
head(eig.val)
```

```{r}
#scree plot showing percentages of inertia explained by each FAMD dimensions
fviz_screeplot(res.famd)
```

```{r}
#list containing the coordinates, the cos2 and the contribution of all variables
var <- get_famd_var(res.famd)
var
# Coordinates of variables
head(var$coord)
# Cos2: quality of representation on the factor map
head(var$cos2)
# Contributions to the  dimensions
head(var$contrib)
```

```{r}
#correlation between variables
#and the principal dimensions
#and the contribution of variables to the dimensions 1 and 2
# Plot of variables
fviz_famd_var(res.famd, repel = TRUE)
# Contribution to the first dimension
fviz_contrib(res.famd, "var", axes = 1)
# Contribution to the second dimension
fviz_contrib(res.famd, "var", axes = 2)
#note: red dashed line is the expected average value if all variables contributed the same
```

```{r}
#quantitative vars
quanti.var <- get_famd_var(res.famd, "quanti.var")
quanti.var 
#plot showing the relationship between variables, 
#the quality of the representation of variables, and 
#the correlation between variables and the dimensions
fviz_famd_var(res.famd, "quanti.var", repel = TRUE,
              col.var = "black")
#plot above but coloring the most contributing quantitative variables
fviz_famd_var(res.famd, "quanti.var", col.var = "contrib", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE)
#plot above but coloring by the variable's cos2 values to represent the quality of 
#representation on the factor map.
#If a variable is well represented by two dimensions, the sum of the cos2 is closed to one.
fviz_famd_var(res.famd, "quanti.var", col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE)
```

## Data Separation

```{r}
# separate data into train and test datasets
set.seed(478)

tsize <- nrow(presskicks1)*0.7
sub <- sample(1:nrow(presskicks1), size = tsize)
train1 <- presskicks1[sub,]
test1_ <- presskicks1[-sub,]
test1 <- test1_ %>% select(-field_goal_result)

tsize <- nrow(presskicks3)*0.7
sub <- sample(1:nrow(presskicks3), size = tsize)
train3 <- presskicks3[sub,]
test3_ <- presskicks3[-sub,]
test3 <- test3_ %>% select(-field_goal_result)

tsize <- nrow(presskicks5)*0.7
sub <- sample(1:nrow(presskicks5), size = tsize)
train5 <- presskicks5[sub,]
test5_ <- presskicks5[-sub,]
test5 <- test5_ %>% select(-field_goal_result)
```

## KNN

```{r}
df_knn <- df
df_knn$field_goal_result[df_knn$field_goal_result=="blocked"] <- "missed"
```

```{r}
#transform all character variables to numeric so they work with KNN
df_knn$field_goal_result <- as.numeric(as.factor(df_knn$field_goal_result))
df_knn$location <- as.numeric(as.factor(df_knn$location))
df_knn$roof <- as.numeric(as.factor(df_knn$roof))
df_knn$surface <- as.numeric(as.factor(df_knn$surface))
df_knn$posteam_type <- as.numeric(as.factor(df_knn$posteam_type))
df_knn <- select_if(df_knn, is.numeric)   
```

```{r}
#split data into training and testing data (90/10)
cutoff <- sample(1:nrow(df_knn), 0.9*nrow(df_knn))
train <- df_knn[cutoff,]
test <- df_knn[-cutoff,]
#function to calculate accuracy
accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}
```

```{r}
#knn
target_cat <- train[,7]
test_cat <- test[,7]
knn13 <- knn(train, test, cl=target_cat,k=13)
confusion13 <-  table(knn13, test_cat)
confusion13
accuracy(confusion13)
#this model has 82.27468% accuracy
```

```{r}
df_knn2 <- subset(df_knn, select = -c(week, sp, posteam_score_post,
                                      defteam_score_post, wp, def_wp, home_wp,
                                      away_wp, no_score_prob, opp_fg_prob, 
                                      score_differential_post))
head(df_knn2)
cutoff <- sample(1:nrow(df_knn2), 0.9*nrow(df_knn2))
train <- df_knn2[cutoff,]
test <- df_knn2[-cutoff,]
target_cat <- train[,5]
test_cat <- test[,5]
knn21 <- knn(train, test, cl=target_cat,k=21)
confusion21 <-  table(knn21, test_cat)
confusion21
accuracy(confusion21)
#this model has 83.17597% accuracy
```

## 1 Minute Analysis

```{r, message=FALSE}
# kick outcomes and correlation of continuous variables
table(presskicks1$field_goal_result)
# 78% made
presskicks1 %>% select(week,game_seconds_remaining,drive,kick_distance,
                       posteam_timeouts_remaining,defteam_timeouts_remaining,
                       score_differential,no_score_prob,opp_fg_prob,wp,result,
                       temp,wind) %>% ggpairs()
```

### Logistic Regression

```{r}
kicks_timeout <- glm(as.factor(field_goal_result) ~ timeout, data = train1, family = binomial)
#:(
ggplot()+ 
  geom_point(aes(predict(kicks_timeout, test1), test1_$field_goal_result))
```

```{r}
kicks_dtimeouts <- glm(as.factor(field_goal_result) ~ defteam_timeouts_remaining, data = train1, family = binomial)
#:( this model sucks
ggplot()+
  geom_point(aes(predict(kicks_dtimeouts, test1), test1_$field_goal_result))
```

```{r}
#model of things that felt helpful to me
kicks_2 <- glm(as.factor(field_goal_result) ~ game_seconds_remaining + drive + defteam_timeouts_remaining + score_differential + no_score_prob + opp_fg_prob + wp, data = train1, family = binomial)
#predict(kicks_2, test1)
ggplot()+ 
  geom_point(aes(predict(kicks_2, test1), test1_$field_goal_result))
```

```{r}
kicks_3 <- glm(as.factor(field_goal_result) ~ no_score_prob + opp_fg_prob, data = train1, family = binomial)
ggplot()+ 
  geom_point(aes(predict(kicks_3, test1), test1_$field_goal_result))
```

```{r}
kicks_pos <- glm(as.factor(field_goal_result) ~ posteam, data = train1, family = binomial)
# BAL, BUF, DET, JAX, LV, NE
ggplot()+ 
  geom_point(aes(predict(kicks_pos, test1), test1_$field_goal_result))
```

```{r}
kicks_secs <- glm(as.factor(field_goal_result) ~ game_seconds_remaining, data = train1, family = binomial)
ggplot()+ 
  geom_point(aes(predict(kicks_secs, test1), test1_$field_goal_result))
```

```{r}
#sometimes this works and sometimes it doesnt.
#kicks_whole <- glm(as.factor(field_goal_result) ~ ., data = train1, family = binomial)
#ggplot()+
  #geom_point(aes(predict(kicks_whole, test1), test1_$field_goal_result))
```

### Mixed Effects Logistic Regression

```{r, warning=FALSE}
model1 <- glmer(factor(field_goal_result) ~ (1|kicker_player_id) +
                  game_seconds_remaining + kick_distance +
                  defteam_timeouts_remaining +
                  score_differential,
                data = train1, family = binomial)
pred <- predict(model1,test1,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test1_$field_goal_result,pred)
```

Kick distance, defensive timeouts remaining, and score differential are significant.

```{r, warning=FALSE}
model2 <- glmer(factor(field_goal_result) ~ (1|kicker_player_id) + season_type +
                  posteam_type + game_seconds_remaining + drive +
                  kick_distance + timeout + posteam_timeouts_remaining +
                  defteam_timeouts_remaining + score_differential + opp_fg_prob +
                  location + result + roof + surface +
                  temp + wind,
                data = train1, family = binomial)
pred <- predict(model2,test1,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test1_$field_goal_result,pred)
```

Kick distance is significant.

```{r}
postypemod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+posteam_type,
                 data = train1, family = binomial)
pred <- predict(postypemod,test1,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test1_$field_goal_result,pred)
```

No evidence for a home field advantage.

```{r}
gamesecsmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       game_seconds_remaining,
                 data = train1, family = binomial)
pred <- predict(gamesecsmod,test1,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test1_$field_goal_result,pred)
summary(gamesecsmod)
```

```{r}
roofmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       roof,
                 data = train1, family = binomial)
pred <- predict(roofmod,test1,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test1_$field_goal_result,pred)
```

```{r}
tempmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       temp,
                 data = train1, family = binomial)
pred <- predict(tempmod,test1,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test1_$field_goal_result,pred)
```

```{r}
windmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       wind,
                 data = train1, family = binomial)
pred <- predict(windmod,test1,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test1_$field_goal_result,pred)
```

```{r}
kickmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                   kick_distance,
                 data = train1, family = binomial)
pred <- predict(kickmod,test1,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test1_$field_goal_result,pred)
```

```{r}
model3 <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                  game_seconds_remaining + roof + temp + wind + kick_distance,
                data = train1, family = binomial)
pred <- predict(model3,test1,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<0] <- 'made'
table(test1_$field_goal_result,pred)
```

### Random Forest

```{r}
train1$field_goal_result <- ifelse(train1$field_goal_result == "made",1,0)
test1_$field_goal_result <- ifelse(test1_$field_goal_result == "made",1,0)
train1 <- train1 %>% select(-timeout_team)
test1_ <- test1_ %>% select(-timeout_team)
```

```{r, message=FALSE, warning=FALSE}
# Boruta alg for RF variable selection 
boruta <- Boruta(as.factor(field_goal_result) ~ .-timeout_team, 
                   data = presskicks1, doTrace = 2, maxRuns = 500)
print(boruta$finalDecision)
par(mar = c(8,5,2,2))
plot(boruta, las = 2, cex.axis = 0.7)
```

Boruta is a feature selection algorithm. Precisely, it works as a wrapper 
algorithm around Random Forest. 

```{r}
# Random Forest w/o redundant sp variable
set.seed(478)
# Train
rforest_import <- randomForest(as.factor(field_goal_result) ~ ., 
                   data = train1, 
                   ntree = 200, importance = TRUE)
rforest_import
vip(rforest_import)
# Test
rforest_import <- randomForest(as.factor(field_goal_result) ~ ., 
                   data = test1_, 
                   ntree = 200, importance = TRUE)
rforest_import
vip(rforest_import)
```

Only timeouts:
Training acc: 67 + 278 / 67 + 278 + 1 + 20 = 0.94262295082 = 94.26%
Testing acc:8 + 129 / 8 + 129 + 1 + 20 = 0.867088607595 = 86.71% 
All kicks:
Training acc: 92 + 484 / 92 + 484 + 4 + 41 = 0.927536 = 92.75%
Testing acc: 40 + 203 / 40 + 203 + 3 + 21 = 0.910112 = 91.01%

## 3 Minute Analysis

```{r, message=FALSE}
# kick outcomes and correlation of continuous variables
table(presskicks3$field_goal_result)
# 82% made
presskicks3 %>% select(week,game_seconds_remaining,drive,kick_distance,
                       posteam_timeouts_remaining,defteam_timeouts_remaining,
                       score_differential,no_score_prob,opp_fg_prob,wp,result,
                       temp,wind) %>% ggpairs()
```

### Logistic Regression

```{r}
kicks_timeout <- glm(as.factor(field_goal_result) ~ timeout, data = train3, family = binomial)
#:(
ggplot()+ 
  geom_point(aes(predict(kicks_timeout, test3), test3_$field_goal_result))
```

```{r}
kicks_dtimeouts <- glm(as.factor(field_goal_result) ~ defteam_timeouts_remaining, data = train3, family = binomial)
#:( this model sucks
ggplot()+
  geom_point(aes(predict(kicks_dtimeouts, test3), test3_$field_goal_result))
```

```{r}
#model of things that felt helpful to me
kicks_2 <- glm(as.factor(field_goal_result) ~ game_seconds_remaining + drive + defteam_timeouts_remaining + score_differential + no_score_prob + opp_fg_prob + wp, data = train3, family = binomial)
#predict(kicks_2, test)
ggplot()+ 
  geom_point(aes(predict(kicks_2, test3), test3_$field_goal_result))
```

```{r}
kicks_3 <- glm(as.factor(field_goal_result) ~ no_score_prob + opp_fg_prob, data = train3, family = binomial)
ggplot()+ 
  geom_point(aes(predict(kicks_3, test3), test3_$field_goal_result))
```

```{r}
kicks_pos <- glm(as.factor(field_goal_result) ~ posteam, data = train3, family = binomial)
# BAL, BUF, DET, JAX, LV, NE
ggplot()+ 
  geom_point(aes(predict(kicks_pos, test3), test3_$field_goal_result))
```

```{r}
kicks_secs <- glm(as.factor(field_goal_result) ~ game_seconds_remaining, data = train3, family = binomial)
ggplot()+ 
  geom_point(aes(predict(kicks_secs, test3), test3_$field_goal_result))
```

```{r}
#sometimes this works and sometimes it doesnt. 
#kicks_whole <- glm(as.factor(field_goal_result) ~ ., data = train3, family = binomial)
#ggplot()+
  #geom_point(aes(predict(kicks_whole, test3), test3_$field_goal_result))
```

### Mixed Effects Logistic Regression

```{r, warning=FALSE}
model1 <- glmer(factor(field_goal_result) ~ (1|kicker_player_id) +
                  game_seconds_remaining + kick_distance +
                  defteam_timeouts_remaining +
                  score_differential,
                data = train3, family = binomial)
pred <- predict(model1,test3,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test3_$field_goal_result,pred)
```

Kick distance is significant.

```{r, warning=FALSE}
model2 <- glmer(factor(field_goal_result) ~ (1|kicker_player_id) + season_type +
                  posteam_type + game_seconds_remaining + drive +
                  kick_distance + timeout + posteam_timeouts_remaining +
                  defteam_timeouts_remaining + score_differential + opp_fg_prob +
                  location + result + roof + surface +
                  temp + wind,
                data = train3, family = binomial)
pred <- predict(model2,test3,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test3_$field_goal_result,pred)
```

Kick distance, temp, and wind are significant.

```{r}
postypemod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+posteam_type,
                 data = train3, family = binomial)
pred <- predict(postypemod,test3,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test3_$field_goal_result,pred)
```

No evidence for a home field advantage.

```{r, warning=FALSE}
gamesecsmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       game_seconds_remaining,
                 data = train3, family = binomial)
pred <- predict(gamesecsmod,test3,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test3_$field_goal_result,pred)
```

Significant but still predicts all kicks made.

```{r, warning=FALSE}
roofmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       roof,
                 data = train3, family = binomial)
pred <- predict(roofmod,test3,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test3_$field_goal_result,pred)
```

```{r, warning=FALSE}
tempmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       temp,
                 data = train3, family = binomial)
pred <- predict(tempmod,test3,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test3_$field_goal_result,pred)
```

```{r}
windmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       wind,
                 data = train3, family = binomial)
pred <- predict(windmod,test3,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test3_$field_goal_result,pred)
```

Significant, predicts 1 kick missed.

```{r}
kickmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                   kick_distance,
                 data = train3, family = binomial)
pred <- predict(kickmod,test3,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test3_$field_goal_result,pred)
```

Significant, prediction accuracy 81.8%, might at well have predicted all kicks made.

```{r, warning=FALSE}
model3 <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                  game_seconds_remaining + roof + temp + wind + kick_distance,
                data = train3, family = binomial)
pred <- predict(model3,test3,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<0] <- 'made'
table(test3_$field_goal_result,pred)
```

Wind and kick distance significant.

### Random Forest

```{r}
train3$field_goal_result <- ifelse(train3$field_goal_result == "made",1,0)
test3_$field_goal_result <- ifelse(test3_$field_goal_result == "made",1,0)
train3 <- train3 %>% select(-timeout_team)
test3_ <- test3_ %>% select(-timeout_team)
```

```{r, message=FALSE, warning=FALSE}
# Boruta alg for RF variable selection - 180 seconds 
boruta <- Boruta(as.factor(field_goal_result) ~ ., 
                   data = presskicks3, doTrace = 2, maxRuns = 500)
print(boruta$finalDecision)
par(mar = c(8,5,2,2))
plot(boruta, las = 2, cex.axis = 0.7)
```


```{r}
## Random Forest - All variables 
set.seed(478)
# Train 
rforest_import <- randomForest(as.factor(field_goal_result) ~ ., 
                   data = train3, 
                   ntree = 200, importance = TRUE)
rforest_import
vip(rforest_import)
# Test
rforest_import <- randomForest(as.factor(field_goal_result) ~ ., 
                   data = test3_, 
                   ntree = 200, importance = TRUE)
rforest_import
vip(rforest_import)
```

Only timeouts:
Training acc: 82 + 500 / 82 + 500 + 42 + 7 = 0.92234548336 = 92.23%
Testing acc: 21 + 218 / 21 + 218 + 32 + 0 = 0.881918819188 = 88.19%
All kicks:
Training acc: 122 + 1051 / 122 + 1051 + 6 + 104 = 0.914263 = 91.43%
Testing acc: 58 + 438 / 58 + 438 + 5 + 49 = 0.901818 = 90.18%

## 5 Minute Analysis

```{r, message=FALSE}
# kick outcomes and correlation of continuous variables
table(presskicks5$field_goal_result)
# 83% made
presskicks5 %>% select(week,game_seconds_remaining,drive,kick_distance,
                       posteam_timeouts_remaining,defteam_timeouts_remaining,
                       score_differential,no_score_prob,opp_fg_prob,wp,result,
                       temp,wind) %>% ggpairs()
```

### Logistic Regression

```{r}
kicks_timeout <- glm(as.factor(field_goal_result) ~ timeout, data = train5, family = binomial)
#:(
ggplot()+ 
  geom_point(aes(predict(kicks_timeout, test5), test5_$field_goal_result))
```

```{r}
kicks_dtimeouts <- glm(as.factor(field_goal_result) ~ defteam_timeouts_remaining, data = train5, family = binomial)
#:( this model sucks
ggplot()+
  geom_point(aes(predict(kicks_dtimeouts, test5), test5_$field_goal_result))
```

```{r}
#model of things that felt helpful to me
kicks_2 <- glm(as.factor(field_goal_result) ~ game_seconds_remaining + drive + defteam_timeouts_remaining + score_differential + no_score_prob + opp_fg_prob + wp, data = train5, family = binomial)
#predict(kicks_2, test)
ggplot()+ 
  geom_point(aes(predict(kicks_2, test5), test5_$field_goal_result))
```

```{r}
kicks_3 <- glm(as.factor(field_goal_result) ~ no_score_prob + opp_fg_prob, data = train5, family = binomial)
ggplot()+ 
  geom_point(aes(predict(kicks_5, test3), test5_$field_goal_result))
```

```{r}
kicks_pos <- glm(as.factor(field_goal_result) ~ posteam, data = train5, family = binomial)
# BAL, BUF, DET, JAX, LV, NE
ggplot()+ 
  geom_point(aes(predict(kicks_pos, test5), test5_$field_goal_result))
```

```{r}
kicks_secs <- glm(as.factor(field_goal_result) ~ game_seconds_remaining, data = train5, family = binomial)
ggplot()+ 
  geom_point(aes(predict(kicks_secs, test5), test5_$field_goal_result))
```

```{r}
#sometimes this works and sometimes it doesnt. 
#kicks_whole <- glm(as.factor(field_goal_result) ~ ., data = train5, family = binomial)
#ggplot()+
  #geom_point(aes(predict(kicks_whole, test5), test5_$field_goal_result))
```

### Mixed Effects Logistic Regression

```{r, warning=FALSE}
model1 <- glmer(factor(field_goal_result) ~ (1|kicker_player_id) +
                  game_seconds_remaining + kick_distance +
                  defteam_timeouts_remaining +
                  score_differential,
                data = train5, family = binomial)
pred <- predict(model1,test5,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test5_$field_goal_result,pred)
```

Kick distance is significant.

```{r, warning=FALSE}
model2 <- glmer(factor(field_goal_result) ~ (1|kicker_player_id) + season_type +
                  posteam_type + game_seconds_remaining + drive +
                  kick_distance + timeout + posteam_timeouts_remaining +
                  defteam_timeouts_remaining + score_differential + opp_fg_prob +
                  location + result + roof + surface +
                  temp + wind,
                data = train5, family = binomial)
pred <- predict(model2,test5,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test5_$field_goal_result,pred)
```

Kick distance is significant.

```{r}
postypemod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+posteam_type,
                 data = train5, family = binomial)
pred <- predict(postypemod,test5,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test5_$field_goal_result,pred)
```

No evidence for a home field advantage.

```{r, warning=FALSE}
gamesecsmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       game_seconds_remaining,
                 data = train5, family = binomial)
pred <- predict(gamesecsmod,test5,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test5_$field_goal_result,pred)
```

Significant but still predicts all kicks made.

```{r, warning=FALSE}
roofmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       roof,
                 data = train5, family = binomial)
pred <- predict(roofmod,test5,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test5_$field_goal_result,pred)
```

```{r, warning=FALSE}
tempmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       temp,
                 data = train5, family = binomial)
pred <- predict(tempmod,test5,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test5_$field_goal_result,pred)
```

```{r}
windmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                       wind,
                 data = train5, family = binomial)
pred <- predict(windmod,test5,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test5_$field_goal_result,pred)
```

```{r}
kickmod <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                   kick_distance,
                 data = train5, family = binomial)
pred <- predict(kickmod,test5,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<=0] <- 'made'
table(test5_$field_goal_result,pred)
```

Significant, prediction accuracy 84%, might at well have predicted all kicks made.

```{r, warning=FALSE}
model3 <- glmer(factor(field_goal_result)~(1|kicker_player_id)+
                  game_seconds_remaining + roof + temp + wind + kick_distance,
                data = train5, family = binomial)
pred <- predict(model3,test5,allow.new.levels=T)
pred[pred>0] <- 'missed'
pred[pred<0] <- 'made'
table(test5_$field_goal_result,pred)
```

Game seconds remaining and kick distance significant.

### Random Forest

```{r}
train5$field_goal_result <- ifelse(train5$field_goal_result == "made",1,0)
test5_$field_goal_result <- ifelse(test5_$field_goal_result == "made",1,0)
train5 <- train5 %>% select(-timeout_team)
test5_ <- test5_ %>% select(-timeout_team)
```

```{r, message=FALSE, warning=FALSE}
# Boruta alg for RF variable selection - 300 seconds 
boruta <- Boruta(as.factor(field_goal_result) ~ ., 
                   data = presskicks5, doTrace = 2, maxRuns = 500)
print(boruta$finalDecision)
par(mar = c(8,5,2,2))
plot(boruta, las = 2, cex.axis = 0.7)
```


```{r}
## Random Forest - All variables 
set.seed(478)
# Train 
rforest_import <- randomForest(as.factor(field_goal_result) ~ ., 
                   data = train5, 
                   ntree = 200, importance = TRUE)
rforest_import
vip(rforest_import)
# Test
rforest_import <- randomForest(as.factor(field_goal_result) ~ ., 
                   data = test5_, 
                   ntree = 200, importance = TRUE)
rforest_import
vip(rforest_import)
```

Only timeouts:
Training acc: 82 + 500 / 82 + 500 + 42 + 7 = 0.92234548336 = 92.23%
Testing acc: 21 + 218 / 21 + 218 + 32 + 0 = 0.881918819188 = 88.19%
All kicks:
Training acc: 150 + 1527 / 150 + 1527 + 7 + 147 = 0.915893 = 91.59%
Testing acc: 59 + 645 / 59 + 645 + 6 + 76 = 0.89567 = 89.57%

